---
title: "CC project"
author: "Jingxi.Feng (ID:jf3670) & Tracy Mao (ID:jm5893)"
date: "2024-11-05"
output: html_document
---
```{r}
library(ggplot2)
library(tidyr) 
library(dplyr)
library(VIM)
library(mice)
```

# Four Methods in here: 1.Remove NA. 2.Replace by Mean/Median/Mod. 3.KNN imputation. 4. Predictive Model 


# Part A).
## Let create a "fake" dataset to see how to handle missing value 
### Assign random missing value 

```{r}
set.seed(42)

df <- data.frame(
  Numerical_1 = sample(1:100, 100, replace = TRUE),
  Numerical_2 = sample(50:150, 100, replace = TRUE),
  Category_1 = sample(c('A', 'B', 'C', 'D'), 100, replace = TRUE),
  Category_2 = sample(c('X', 'Y', 'Z'), 100, replace = TRUE)
)

df_none_missing <- df
```




### Before moving next let see its distribution
```{r}
par(mfrow = c(2, 2))

# Loop through each column of the dataframe
for (col_name in names(df)) {
  
  if (is.numeric(df[[col_name]])) {
    hist(df[[col_name]], 
         main = paste("Distribution of", col_name), 
         xlab = col_name, 
         col = "skyblue", 
         border = "white")
    
  } else {
    barplot(table(df[[col_name]]), 
            main = paste("Distribution of", col_name), 
            xlab = col_name, 
            col = "coral", 
            border = "white")
  }
}

par(mfrow = c(1, 1))
```


```{r}
# Random choose 2 missing in num column 1
missing_indices_num1 <- sample(1:100, 2)

# Random choose 10 missing in num column 2
missing_indices_num2 <- sample(1:100, 10)

# Random choose 25 missing in cat column 1
missing_indices_cat1 <- sample(1:100, 25)

# Random choose 5 missing in cat column 2
missing_indices_cat2 <- sample(1:100, 5)


# Assign the missing value to the previous select columns
df$Numerical_1[missing_indices_num1] <- NA
df$Numerical_2[missing_indices_num2] <- NA
df$Category_1[missing_indices_cat1] <- NA
df$Category_2[missing_indices_cat2] <- NA


```


### Visualize the missing values
```{r}

## These just all preprocessing

df_missing <- df %>%
  mutate(across(everything(), as.character)) %>%  # Convert all columns to character
  mutate(row_id = row_number()) %>%
  pivot_longer(cols = -row_id, names_to = "variable", values_to = "value")

# Mark missing values
df_missing <- df_missing %>%
  mutate(is_missing = is.na(value))

# Plot the missing values
ggplot(df_missing, aes(x = variable, y = row_id, fill = is_missing)) +
  geom_tile(color = "white") +
  scale_fill_manual(values = c("FALSE" = "blue", "TRUE" = "red")) +
  labs(title = "Missing Values Heatmap",
       x = "Variables",
       y = "Row Number",
       fill = "Missing") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# B). We could start our work for handling missing value 


## 1.a). Remove the missing value by row 
```{r}

df_clean_remove_rows <- na.omit(df)
head(df_clean_remove_rows)

```

## 1.b). visualize after method 1 
```{r}
par(mfrow = c(2, 4))  

for (col_name in names(df_none_missing)) {
  
  # Plot original distribution
  if (is.numeric(df_none_missing[[col_name]])) {
    hist(df_none_missing[[col_name]], 
         main = paste("Original", col_name), 
         xlab = col_name, 
         col = "skyblue", 
         border = "white")
    
  } else {
    barplot(table(df_none_missing[[col_name]]), 
            main = paste("Original", col_name), 
            xlab = col_name, 
            col = "coral", 
            border = "white")
  }
  
  # Plot modified distribution after removing rows with missing values
  if (is.numeric(df_clean_remove_rows[[col_name]])) {
    hist(df_clean_remove_rows[[col_name]], 
         main = paste("Removed rows", col_name), 
         xlab = col_name, 
         col = "lightgreen", 
         border = "white")
    
  } else {
    barplot(table(df_clean_remove_rows[[col_name]]), 
            main = paste("Removed rows", col_name), 
            xlab = col_name, 
            col = "lightpink", 
            border = "white")
  }
}

# Reset plotting area to default
par(mfrow = c(1, 1))

## Notice the scale of new one now become smaller 

```



## 2.a). Mutate missing with Mean, Median,or Mode

```{r}
# 1. Mean Imputation for Numerical_1
df_mean_impute <- df |>
  mutate(Numerical_1 = ifelse(is.na(Numerical_1), 
                              mean(Numerical_1, na.rm = TRUE), 
                              Numerical_1))

# 2. Median Imputation for Numerical_2
df_median_impute <- df_mean_impute |>
  mutate(Numerical_2 = ifelse(is.na(Numerical_2), 
                              median(Numerical_2, na.rm = TRUE), 
                              Numerical_2))

# 3. Mode Imputation for Categorical Variables
mode_imputation <- function(x) {
  ux <- unique(x)
  ux <- ux[!is.na(ux)]  # Exclude NA from the unique values
  ux[which.max(tabulate(match(x, ux)))]
}

df_mode_impute <- df_median_impute |>
  mutate(Category_1 = ifelse(is.na(Category_1), 
                             mode_imputation(Category_1), 
                             Category_1),
         Category_2 = ifelse(is.na(Category_2), 
                             mode_imputation(Category_2), 
                             Category_2))
```





## 2.b). Visualize the method 2
```{r}


# New plots with comparison
par(mfrow = c(2, 4))  # Set up a 2x4 grid for side-by-side comparison

for (col_name in names(df)) {
  
  # Plot original distribution
  if (is.numeric(df_none_missing[[col_name]])) {
    hist(df_none_missing[[col_name]], 
         main = paste("Original", col_name), 
         xlab = col_name, 
         col = "skyblue", 
         border = "white")
    
  } else {
    barplot(table(df_none_missing[[col_name]]), 
            main = paste("Original", col_name), 
            xlab = col_name, 
            col = "coral", 
            border = "white")
  }
  
  # Plot modified distribution
  if (is.numeric(df_mode_impute[[col_name]])) {
    hist(df_mode_impute[[col_name]], 
         main = paste("Mean/Median", col_name), 
         xlab = col_name, 
         col = "lightgreen", 
         border = "white")
    
  } else {
    barplot(table(df_mode_impute[[col_name]]), 
            main = paste("Mode", col_name), 
            xlab = col_name, 
            col = "lightpink", 
            border = "white")
  }
}

# Reset plotting area to default
par(mfrow = c(1, 1))


## Notice now the one would have more "normal" and center-clustering 
```



## 3.a) Using KNN to replace the Missing by refering the "near" data point 
```{r}


df_knn_impute <- kNN(df, k = 5, imp_var = FALSE)
head(df_knn_impute)

```

## 3.b) Visualize from method 3
```{r}
par(mfrow = c(2, 4))  

for (col_name in names(df)) {
  
  # Plot original distribution
  if (is.numeric(df_none_missing[[col_name]])) {
    hist(df_none_missing[[col_name]], 
         main = paste("Original", col_name), 
         xlab = col_name, 
         col = "skyblue", 
         border = "white")
    
  } else {
    barplot(table(df_none_missing[[col_name]]), 
            main = paste("Original", col_name), 
            xlab = col_name, 
            col = "coral", 
            border = "white")
  }
  
  # Plot modified distribution after KNN imputation
  if (is.numeric(df_knn_impute[[col_name]])) {
    hist(df_knn_impute[[col_name]], 
         main = paste("KNN Imputed", col_name), 
         xlab = col_name, 
         col = "lightgreen", 
         border = "white")
    
  } else {
    barplot(table(df_knn_impute[[col_name]]), 
            main = paste("KNN Imputed", col_name), 
            xlab = col_name, 
            col = "lightpink", 
            border = "white")
  }
}


par(mfrow = c(1, 1))



## Notice in here, the KNN is create a more "close" to true/orignial data when missing is small
## And even missing is large, it does not necessary different so much from the true one
```


## 4.a). Predictive Model to handle missing value 
```{r}
# Step 1: Copy the original dataset

# Step 1: Load or create your dataset
  # Load a dataset named 'df' which contains missing values
# Step 2: Convert categorical columns to factors
  # Ensure categorical columns are recognized as factors for appropriate handling during imputation
# Step 3: Define the imputation methods for each column
  # Specify imputation methods for each column:
  # 'pmm' (Predictive Mean Matching) for numerical columns
  # 'polyreg' (Polytomous Logistic Regression) for categorical columns
# Step 4: Apply MICE to the dataset
  # Impute missing values using the MICE package
# Step 5: Extract the completed dataset
  # Retrieve the imputed dataset from the MICE object


# Use mice function

 


# df <- your_dataset  

#df$Category_1 <- as.factor(df$Category_1)  # Convert 'Category_1' to a factor

#df$Category_2 <- as.factor(df$Category_2)  # Convert 'Category_2' to a factor

# methods <- c("pmm", "pmm", "polyreg", "polyreg") 

# mice_imputed <- mice(df, m = 1, method = methods, seed = 42)


# df_imputed <- complete(mice_imputed)



```








